---
title: "Infectious Disease Sample Report"
date: "`r Sys.Date()`"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(readr)
library(sf)
library(dplyr)
library(RColorBrewer)
```

# Executive Summary
This report presents an analysis of infectious disease surveillance in the Philippines. It includes total cases by disease and weekly trends.

```{r}
# Load and Prepare Data
# Simulated example: Replace this with your actual DOH dataset CSV
# Example: disease_data <- read_csv("doh_weekly_cases.csv")
set.seed(123)

# Simulate weekly data for demonstration
diseases <- c("Dengue", "Measles")

date_seq <- seq(as.Date("2024-01-01"), as.Date("2024-12-31"), by = "week")

disease_data <- expand.grid(
  date = date_seq,
  disease = diseases
)

disease_data$cases <- sample(0:100, nrow(disease_data), replace = TRUE)

# Create a weekly summary
disease_data_weekly <- disease_data %>%
  group_by(week = floor_date(date, "week"), disease) %>%
  summarise(total_cases = sum(cases), .groups = "drop")
```

# Text Summary
```{r}
# Calculate the most reported disease
top_disease <- disease_data_weekly %>%
  group_by(disease) %>%
  summarise(total = sum(total_cases)) %>%
  arrange(desc(total)) %>%
  slice(1)

# Find the peak week
peak_week <- disease_data_weekly %>%
  arrange(desc(total_cases)) %>%
  slice(1)
```
As of `r Sys.Date()`, the most reported disease is `r top_disease$disease` with a total of `r format(top_disease$total, big.mark = ",")` cases.

The peak week was the week of `r format(peak_week$week, "%B %d, %Y")`, when `r format(peak_week$total_cases, big.mark = ",")` cases of `r peak_week$disease` were reported.

# Bar Chart: Total Cases by Disease
```{r}
# Total cases by disease
total_cases_by_disease <- disease_data %>%
  group_by(disease) %>%
  summarise(total = sum(cases))

# Plot the bar chart
ggplot(total_cases_by_disease, aes(x = disease, y = total, fill = disease)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Reported Cases by Disease", y = "Total Cases", x = "") +
  theme_minimal()
```

# Time Series Plot: Weekly Trends
```{r}
# Weekly trends of cases
ggplot(disease_data_weekly, aes(x = week, y = total_cases, color = disease)) +
  geom_line() +
  geom_point() +
  labs(title = "Weekly Reported Cases of Infectious Diseases",
       x = "Week",
       y = "Total Cases") +
  theme_minimal() +
  scale_x_date(date_labels = "%b %Y", date_breaks = "1 month") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```